Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-06-24 kappa.tex
Document basal sliding.

\begin{equation}
    \bm{\tau}_b = -\tau_c \frac{\vec{v}_b}
                               {{u_{th}}^q\,|\vec{v}_b|^{1-q}}
\end{equation}

\begin{equation}
   \tau_c = c_{0} + \tan{\phi}\cdot N_{til}
\end{equation}

\newcommand{\phimin}{\phi_{\mathrm{min}}}
\newcommand{\phimax}{\phi_{\mathrm{max}}}
\newcommand{\bmin}{b_{\mathrm{min}}}
\newcommand{\bmax}{b_{\mathrm{max}}}

\begin{equation}
  \phi(x,y) =
  \begin{cases}
    \phimin, & b(x,y) \le \bmin, \\
    \phimin + (b(x,y) - \bmin) \,M, & \bmin < b(x,y) < \bmax, \\
    \phimax, & \bmax \le b(x,y).
  \end{cases}
  \label{eq:phipiecewise}
\end{equation}

\begin{equation}
N_{til} = \delta P_o \, 10^{(e_0/C_c)
          \left(1 - (W_{til}/W_{til}^{max})\right)}
\end{equation}

>>>>

Although the SIA is valid only under non-sliding conditions, the SSA requires
a sliding law as basal boundary condition, which in fact constitutes the main
control on SSA velocities and in turn locations of ice streams.

The simulations presented in this thesis use a pseudo-plastic sliding law,

\begin{equation}
    \label{eqn:pseudoplastic}
    \vec{\tau}_b = -\tau_c \frac{\vec{v}_b}{{v_{th}}^q\,|\vec{v}_b|^{1-q}},
\end{equation}

where $\vec{\tau}_b$ is the basal traction force, $\vec{v}_b$ is the basal
velocity, and $v_{th}$ is an arbitrary velocity threshold. In the case of
$q=0$, equation~\ref{eqn:pseudoplastic} corresponds to a purely plastic law,

\begin{equation}
    \label{eqn:plastic}
    \vec{\tau}_b = -\tau_c \frac{\vec{v}_b}{|\vec{v}_b|}.
\end{equation}

However a non-zero exponent is chosen here in order to improve convergence. The
yield stress $\tau_c$ is related to till properties by the Mohr-Coulomb
criterion,

\begin{equation}
   \tau_c = c_0 + N\,\tan{\phi}.
\end{equation}

The till cohesion $c_0$ is assumed to be zero. The effective pressure on the
till, $N$, is determined by the modelled amount of water at the bed. However,
two different parametrisations are used within this thesis. In Paper~I, using
PISM~0.5, the effective pressure is linearly related to water content by a
simple parametrisation,

\begin{equation}
    N = \rho gH (1 - \alpha \frac{W}{W_{max}}).
\end{equation}

In Paper~IV, using PISM~0.6, the effective pressure is determined by an
empirical parametrisation based on laboratory experiments on Antarctic till,

\begin{equation}
    N = \delta \rho gH \, 10^{(e_0/C_c) (1 - (W/W_{max}))}
\end{equation}

Drainage of water at the base of the ice sheet, although implemented in PISM,
is not included in simulations presented here for the sake of simplicity.
In turn, the amount of water at the bed, $W$, corresponds to the local
accumulation of basal melt-water. In varies from zero to $W_{max}$, a threshold
above which further melt exits the system.

Finally, the till friction angle $\phi$ is taken a~function of modern bed
elevation, with lowest values occurring at low elevations, thus accounting for
a weakening of till associated with the presence of marine sediments,

\begin{equation}
  \phi(x,y) =
  \begin{cases}
    \phi_0, & \mathrm{for}\ b \le b_0, \\
    \phi_0 + (\phi_1-\phi_0) \frac{b - b_0}{b_1-b_0},
        & \mathrm{for}\ b_0 < b < b_1, \\
    \phi_1, & \mathrm{for}\ b_1 \le b.
  \end{cases}
  \label{eq:phipiecewise}
\end{equation}

where $b_0=0$\,m is modern sea-level and $b_1=200$\,m is a rough average of
highest shoreline elevations in the model domain. The values of $\phi_1$ and
$\phi_2$ are chosen differently in Paper~I (10 and 30$^\circ$) and Paper~IV
(15 and 45$^\circ$), but these choices are roughly equivalent in the case of a
saturated till ($W=W_{max}$).
